### 2016-06-26 ch01-intro/introduction.tex
Small changes (CH01, some CH04)

\section{Modeling}
Models play an important role in today's world, even though we often are not always aware of them.
In particular, models enable us to predict the future (within a certain bound of uncertainty):
\begin{itemize}
  \item supermarkets try to predict how much fruit and vegetables their regular customers buy such that they can order a large enough supply,
  \item traders on the stock market aim to predict whether companies such as Tesla or Apple will continue to prosper over the next quarter such that they can decide if it's time to buy more stock or rather sell,
  \item weather forecasters hope to predict what weather it might become tomorrow,
  \item how much electricity will the whole country use next week (do we need to start up a new power plant or not),
  \item how much noise the bodywork of a car will produce when it is driving over a bumpy road,
  \item what the exhaust of nitrogen oxides of a car will be for a certain operating point of a diesel engine, 
  \item are the vibrations in a robot are small enough for it to carry out very precise tasks such as positioning electronic components with sub-millimeter precision,
\end{itemize}
and so on.

@@

\paragraph{White box}
On the one hand, there is so-called \term{`white box'} modeling, or first-principles modeling.
The model is constructed by writing down the equations that govern the system, e.g. Newton's laws of motion for mechanical systems, Kirchhoff's laws for electrical circuits, \latin{etc.}
This is the kind of modeling that is practiced in school, which explains why it is also popular in industry.
Unfortunately, white box modeling quickly becomes unwieldy once more complicated systems are to be studied or many parts of the system are uncertain.
In general, lot of man-hours are required to build a white box model, such that this is often prohibitively expensive for engineering applications.

\paragraph{Black box}
On the other hand, the diametrically opposed approach is called \term{`black box'} modeling.
Here, the model structure is chosen such that a wide variety of phenomena can be described.
Regardless of the nature of the underlying system, be it mechanical, chemical, electrical, structural, \ldots; most dynamic phenomena can be described using differential equations (or difference equations).
In particular, for black-box models, the structure of the model is chosen such that the data is well-approximated.
The downside of this is that interpreting the model physically is not always possible.
Fortunately, the goal of a model is often not to interpret the physical reality but to approximate it well enough to allow engineering

@@

\paragraph{Select a model structure}
Once an experiment has been performed, the obtained input-output data are to be processed.
In many cases, the eventual goal of identification is to obtain a parametric model, so the user has to select a model structure and complexity based on both his prior knowledge and on the data.
A useful tool to help choose the model structure and/or model orders, is to inspect a nonparametric model of the system.
One such tool is the frequency response function.
For instance, by inspecting a frequency response function, a control engineer can have a rough idea of the system behavior in term of  approximate resonance frequencies, asymptotic slopes, \latin{etc.}
This gives a qualitative view of the minimally required complexity of the model and is hence a useful diagnostic tool.
However, since most nonparametric estimators enforce very little prior knowledge about the system, noise in the measurements can sometimes be a problem to distinguish the actual system features.

@@

\paragraph{Validate the model}
Finally, the estimated model is validated.
Essentially, one checks whether the obtained model meets the predefined requirements, e.g. with respect to allowable uncertainty, compexity, \ldots
This can happen in a few different ways.
On the one hand, in-sample measures such as statistical tests on the value of the cost function, correlation tests on the residuals, \latin{etc.} allow to verify that the model performs well on the measured dataset.
Such approaches help to detect underfitting, where the model is not flexible enough to describe the measurements.
On the other hand, the ultimate test is to validate the model against new measurements that have not been used during the estimation.
This prevents overfitting of the model, where the model not only captures the behavior of the system but also describes the particular realization of the noise in the measurements.
The latter leads to models that are overly tuned to the measurements such that their results cannot be replicated for new measurements.
In case the model is not validated, this provides the user with new knowledge: that particular model is not good enough and at least one of the previous steps (or the prior knowledge) has to be altered.

@@

The non-parametric \gls{FRF} estimation methods in \chapref{sec:nonparametric} are based on a yet-unpublished manuscript.
The so-called time-truncated \gls{LPM} that is presented in the same chapter is based on a journal published in \gls{IEEE} Transactions on Instrumentation \& Measurement:
\begin{itemize}
  \item \fullcite{Lumori2014TIM}.
\end{itemize}
In particular, this smoother enables one to reduce the effect of noise on an estimated \gls{FRF} using an automated approach.
Relatedly, a preliminary study of the \gls{LPM} in the context of lightly-damped \gls{MIMO} systems has been presented at the (non-refereed) Benelux Meeting on Systems and Control:
\begin{itemize}
    \item \fullcite{Verbeke2015Benelux}.
\end{itemize}

The use of the non-parametric \gls{FRF} estimation methods for \Hinf gain estimation (and in general the \gls{FRF} interpolation from \chapref{sec:hinf}) has been presented at the 2014 \gls{IFAC} World Conference in South Africa and the 2014 Leuven Conference on Noise and Vibration Engineering (ISMA) in Leuven:
\begin{itemize}
  \item \fullcite{Geerardyn2014IFAC},
  \item \fullcite{Geerardyn2014ISMA}.
\end{itemize}
Preliminary results have been presented at local (non-refereed) conferences:
\begin{itemize}
  \item \fullcite{Geerardyn2013Benelux},
  \item \fullcite{Geerardyn2013ERNSI},
  \item \fullcite{Geerardyn2014Benelux},
  \item \fullcite{Geerardyn2014DYSCO}, and
  \item \fullcite{Geerardyn2014ERNSI}.
\end{itemize}
Experimental work related to \chapref{sec:hinf} has also been presented at the 2015 \gls{IFAC} Symposium on System Identification (\textsc{sysid}) and :
\begin{itemize}
  \item \fullcite{Voorhoeve2015SYSID}
  \item \fullcite{Voorhoeve2015ERNSI}
\end{itemize}

>>>>

\section{Modeling}
Models play an important role in today's world, even though we often are not always aware of them.
In particular, models enable us to predict the future (within a certain bound of uncertainty):
\begin{itemize}
  \item supermarkets try to predict how much fruit and vegetables their regular customers buy such that they can order a large enough supply,
  \item traders on the stock market aim to predict whether companies such as Tesla or Apple will continue to prosper over the next quarter such that they can decide if it's time to buy more stock or rather sell,
  \item weather forecasters hope to predict what weather it might become tomorrow,
  \item how much electricity will the whole country use next week (do we need to start up a new power plant or not),
  \item how much noise the bodywork of a car will produce when it is driving over a bumpy road,
  \item how many nitrogen oxides will a car exhaust for a certain operating point of its diesel engine, 
  \item the vibrations in a robot are small enough for it to carry out very precise tasks such as positioning electronic components with sub-millimeter precision,
\end{itemize}
and so on.

@@

\paragraph{White box}
On the one hand, there is so-called \term{`white box'} modeling, or first-principles modeling.
The model is constructed by writing down the equations that govern the system, e.g. Newton's laws of motion for mechanical systems, Kirchhoff's laws for electrical circuits, \latin{etc.}
This is the kind of modeling that is practiced in school, which explains why it is also popular in industry.
Unfortunately, white box modeling quickly becomes unwieldy once more complicated systems are to be studied or many parts of the system are uncertain.
In general, many man-hours are required to build a white box model, such that this is often prohibitively expensive for engineering applications.

\paragraph{Black box}
On the other hand, the diametrically opposed approach is called \term{`black box'} modeling.
Here, the model structure is chosen such that a wide variety of phenomena can be described.
Regardless of the nature of the underlying system, be it mechanical, chemical, electrical, structural, \ldots; most dynamic phenomena can be described using differential equations (or difference equations).
In particular, for black-box models, the structure of the model is chosen such that the data is well-approximated.
The downside of this is that interpreting the model physically is not always possible.
Fortunately, the goal of a model is often not to interpret the physical reality but to approximate it well enough to allow engineering.

@@

\paragraph{Select a model structure}
Once an experiment has been performed, the obtained input-output data are to be processed.
In many cases, the eventual goal of identification is to obtain a parametric model, so the user has to select a model structure and complexity based on both his prior knowledge and on the data.
A useful tool to help choose the model structure and/or model orders, is to inspect a nonparametric model of the system.
One such tool is the frequency response function.
For instance, by inspecting a frequency response function, a control engineer can have a rough idea of the system behavior in terms of  approximate resonance frequencies, asymptotic slopes, \latin{etc.}
This gives a qualitative view of the minimally required complexity of the model and is hence a useful diagnostic tool.
However, since most nonparametric estimators enforce very little prior knowledge about the system, it can sometimes be problematic to clearly distinguish noise in the measurements and the actual systematic behavior.

@@

\paragraph{Validate the model}
Finally, the estimated model is validated.
Essentially, one checks whether the obtained model meets the predefined requirements, e.g. with respect to allowable uncertainty, complexity, \ldots
This can happen in a few different ways.
On the one hand, in-sample measures such as statistical tests on the value of the cost function, correlation tests on the residuals, \latin{etc.} allow to verify that the model performs well on the measured dataset.
Such approaches help to detect underfitting, where the model is not flexible enough to describe the measurements.
On the other hand, the ultimate test is to validate the model against new measurements that have not been used during the estimation.
This prevents overfitting of the model, where the model not only captures the behavior of the system but also describes the particular realization of the noise in the measurements.
The latter leads to models that are overly tuned to the measurements such that their results cannot be replicated for new measurements.
In case the model is not validated, this provides the user with new knowledge: that particular model is not good enough and at least one of the previous steps (or the prior knowledge) has to be altered.

@@

The non-parametric \gls{FRF} estimation methods in \chapref{sec:nonparametric} are based on a yet-unpublished manuscript.
The so-called time-truncated \gls{LPM} that is presented in the same chapter is based on a journal  article published in \gls{IEEE} Transactions on Instrumentation \& Measurement:
\begin{itemize}
  \item \fullcite{Lumori2014TIM}.
\end{itemize}
In particular, this smoother enables one to reduce the effect of noise on an estimated \gls{FRF} using an automated approach.
Relatedly, a preliminary study of the \gls{LPM} in the context of lightly-damped \gls{MIMO} systems has been presented at the (non-refereed) Benelux Meeting on Systems and Control:
\begin{itemize}
    \item \fullcite{Verbeke2015Benelux}.
\end{itemize}

The use of the non-parametric \gls{FRF} estimation methods for \Hinf gain estimation (and in general the \gls{FRF} interpolation from \chapref{sec:hinf}) has been presented at the 2014 \gls{IFAC} World Congress in South Africa and the 2014 Leuven Conference on Noise and Vibration Engineering (ISMA) in Leuven:
\begin{itemize}
  \item \fullcite{Geerardyn2014IFAC},
  \item \fullcite{Geerardyn2014ISMA}.
\end{itemize}
Preliminary results have been presented at local (non-refereed) conferences:
\begin{itemize}
  \item \fullcite{Geerardyn2013Benelux},
  \item \fullcite{Geerardyn2013ERNSI},
  \item \fullcite{Geerardyn2014Benelux},
  \item \fullcite{Geerardyn2014DYSCO}, and
  \item \fullcite{Geerardyn2014ERNSI}.
\end{itemize}
Experimental work related to \chapref{sec:hinf} has also been presented at the 2015 \gls{IFAC} Symposium on System Identification (SYSID) and at the 2015 workshop of the \gls{ERNSI}:
\begin{itemize}
  \item \fullcite{Voorhoeve2015SYSID}
  \item \fullcite{Voorhoeve2015ERNSI}
\end{itemize}

### 2016-06-26 ch04-lrmhinf/lrmhinf.tex
Small changes (CH01, some CH04)

\subsection{Set-Up}
Consider an output error \gls{LTI} discrete-time \gls{SISO} system $\Delta(q)$ excited by an input signal $u_{\Delta}(n)$.
For an infinitely long data record, this can be described in the time-domain as
\begin{equation}
  y_{\Delta}(n) 
    = 
      y_{\Delta0}(n)
      + v(n) 
    = 
      \Delta(q)u_{\Delta}(n) 
      + H(q)e(n)
  \label{eq:lrmhinf:sysDeltaInfData}
\end{equation}
where $q^{-1}$ is the lag operator, $e(n)$ is white Gaussian noise with unity variance such that $v(n) = H(q)e(n)$ can be colored noise, and both $\Delta(q)$ and $H(q)$ are stable causal real-rational functions.
For a limited data record ($n \in \Set{0,\ldots,N-1}$), the transient contributions $t_{\Delta}(n)$ lead to:
\begin{equation}
    y_{\Delta}(n) 
      = 
        \Delta(q)u(n) 
        + H(q)e(n) 
        + t_{\Delta}(n)
    \text{.}
    \label{eq:lrmhinf:sysDeltaFiniteData}
\end{equation}
By applying the \gls{DFT}
\begin{equation}
  X(k) 
       = 
         \frac{1}
              {\sqrt{N}} 
         \sum_{n=0}^{N-1} 
            x(n) \exp \left( \frac{- j 2 \pi k n }{N} \right)
         \quad   
         \forall k \in \mathbb{Z}
  \label{eq:lrmhinf:DFT}
\end{equation} 
to both sides of~\eqref{eq:lrmhinf:sysDeltaFiniteData}, one obtains:
\begin{equation}
  Y_{\Delta}(k) 
                = 
                  \Delta(\omega_k) U_{\Delta}(k) 
                  + T_{\Delta}(\omega_k) 
                  + V(k) 
  \text{.}
  \label{eq:lrmhinf:sysDeltaFiniteDataFD}
\end{equation}
The $T_{\Delta}$ contribution models the difference between the begin and end conditions of the filters $\Delta$ and $H$ in the measurement~\citep{Pintelon1997ARB}.
The index $k$ corresponds to the $k^{\text{th}}$ frequency bin with frequency $\omega_{k} = 2\pi k / (N \Ts)$ such that the frequency grid $\DFTGrid \isdef \Set{\left.\omega_k \right| k = 1,\ldots,N-1}$.

>>>>

\subsection{Set-Up}
Consider an output error \gls{LTI} discrete-time \gls{SISO} system $\Delta(q)$ excited by an input signal $u_{\Delta}(n)$.
This can be described in the time-domain as
\begin{equation}
  y_{\Delta}(n) 
    = 
      y_{\Delta0}(n)
      + v(n) 
    = 
      \Delta(q)u_{\Delta}(n) 
      + H(q)e(n)
  \label{eq:lrmhinf:sysDeltaInfData}
\end{equation}
where $q^{-1}$ is the lag operator, $e(n)$ is white Gaussian noise with unity variance such that $v(n) = H(q)e(n)$ can be colored noise, and both $\Delta(q)$ and $H(q)$ are stable causal real-rational functions.
For a limited data record ($n \in \Set{0,\ldots,N-1}$ with $N$ the number of data points), the transient contributions $t_{\Delta}(n)$ lead to:
\begin{equation}
    y_{\Delta}(n) 
      = 
        \Delta(q)u(n) 
        + H(q)e(n) 
        + t_{\Delta}(n)
    \text{.}
    \label{eq:lrmhinf:sysDeltaFiniteData}
\end{equation}
By applying the \gls{DFT}
\begin{equation}
  X(k) 
       = 
         \frac{1}
              {\sqrt{N}} 
         \sum_{n=0}^{N-1} 
            x(n) \exp \left( \frac{- j 2 \pi k n }{N} \right)
         \quad   
         \forall k \in \mathbb{Z}
  \label{eq:lrmhinf:DFT}
\end{equation} 
to both sides of~\eqref{eq:lrmhinf:sysDeltaFiniteData}, one obtains:
\begin{equation}
  Y_{\Delta}(k) 
                = 
                  \Delta(\omega_k) U_{\Delta}(k) 
                  + T_{\Delta}(\omega_k) 
                  + V(k) 
  \text{.}
  \label{eq:lrmhinf:sysDeltaFiniteDataFD}
\end{equation}
The $T_{\Delta}$ contribution models the difference between the begin and end conditions of the filters $\Delta$ and $H$ in the measurement~\citep{Pintelon1997ARB}.
The index $k$ corresponds to the $k^{\text{th}}$ frequency bin with frequency $\omega_{k} = 2\pi k / (N \Ts)$ such that the frequency grid $\DFTGrid \isdef \Set{\left.\omega_k \right| k = 1,\ldots,N-1}$.

### 2016-06-26 config/glossaries.inc.tex
Some common commands introduced
Some notes Tomas McKelvey
More entries for list of symbols
Small changes (CH01, some CH04)

\newsymbolglos{A1-SetN}{\NaturalNumbers}{Set of the natural numbers}
\newsymbolglos{A2-SetZ}{\IntegerNumbers}{Set of integer numbers}
\newsymbolglos{A3-SetQ}{\RationalNumbers}{Set of the rational numbers}
\newsymbolglos{A4-SetR}{\RealNumbers}{Set of the real numbers}
\newsymbolglos{A5-SetC}{\ComplexNumbers}{Set of the complex numbers}
\newsymbolglos{A6-SetNnozero}{\NaturalNumbersWithoutZero}{Set of the natural numbers without the number $0$}
\newsymbolglos{A7-SetRnozero}{\RealNumbersWithoutZero}{Set of the positive real numbers without the number $0$, i.e. $\RealNumbers \without \set{0}$}
\newsymbolglos{A8-SetRpos}{\PositiveRealNumbers}{Set of the positive real numbers}
\newsymbolglos{A9-SetRneg}{\NegativeRealNumbers}{Set of negative real numbers}

\newsymbolglos{B-imagunit}{j}{The imaginary unit, $j = \sqrt{-1}$}
\newsymbolglos{B-w}{\omega}{Frequency ($\unit{Hz}$) or angular frequency ($\unit{rad/s}$)}
% \newsymbolglos{B-epsilon}{\eps}{The numerical precision of a floating point number (\code{eps}), e.g. $\eps \approx 2.2 \cdot 10^{-16}$ for \code{double} precision}
\newsymbolglos{B-sys-damping}{\damping}{Relative damping of a pole or system}
\newsymbolglos{B-sys-resonance}{\wn}{Natural (resonance) frequency of a pole or system}
\newsymbolglos{B-sys-timeconst}{\timeconst}{Time constant of a pole or system ($\timeconst \isdef (\damping \wn)^{-1}$)}
\newsymbolglos{B-sys-shiftoperator}{q}{Forward shift operator ($qu(t) = u(t+1)$)}
\newsymbolglos{B-sys-laplaceoperator}{s}{Laplace variable}

\newsymbolglos{count-mc}{n_{\mathrm{MC}}}{Number of Monte Carlo runs}
\newsymbolglos{count-theta}{n_{\theta}}{Number of parameters}
\newsymbolglos{count-window}{\nWind}{Bandwidth of local window}
\newsymbolglos{count-order}{\order{A},\order{B},\order{C}}{Order of local models $\LocalModel{A}, \LocalModel{B},\LocalModel{C}$}

@@

\newacronym{EIV}{EIV}{errors-in-variables}
\newacronym{ETFE}{ETFE}{empirical transfer function estimate}

>>>>

\newsymbolglos{A1-SetN}{\NaturalNumbers}{Set of the natural numbers}
\newsymbolglos{A2-SetZ}{\IntegerNumbers}{Set of integer numbers}
\newsymbolglos{A3-SetQ}{\RationalNumbers}{Set of the rational numbers}
\newsymbolglos{A4-SetR}{\RealNumbers}{Set of the real numbers}
\newsymbolglos{A5-SetC}{\ComplexNumbers}{Set of the complex numbers}
\newsymbolglos{A6-SetNnozero}{\NaturalNumbersWithoutZero}{Set of the natural numbers without the number $0$}
\newsymbolglos{A7-SetRnozero}{\RealNumbersWithoutZero}{Set of the real numbers without the number $0$, i.e. $\RealNumbers \without \set{0}$}
\newsymbolglos{A8-SetRpos}{\PositiveRealNumbers}{Set of the positive real numbers}
\newsymbolglos{A9-SetRneg}{\NegativeRealNumbers}{Set of negative real numbers}

\newsymbolglos{B-imagunit}{j}{The imaginary unit, $j = \sqrt{-1}$}
\newsymbolglos{B-w}{\omega}{Frequency ($\unit{Hz}$) or angular frequency ($\unit{rad/s}$)}
% \newsymbolglos{B-epsilon}{\eps}{The numerical precision of a floating point number (\code{eps}), e.g. $\eps \approx 2.2 \cdot 10^{-16}$ for \code{double} precision}
\newsymbolglos{B-sys-damping}{\damping}{Relative damping of a pole or system}
\newsymbolglos{B-sys-resonance}{\wn}{Natural (resonance) frequency of a pole or system}
\newsymbolglos{B-sys-timeconst}{\timeconst}{Time constant of a pole or system ($\timeconst \isdef (\damping \wn)^{-1}$)}
\newsymbolglos{B-sys-shiftoperator}{q}{Forward shift operator ($q u(t) = u(t+1)$)}
\newsymbolglos{B-sys-laplaceoperator}{s}{Laplace variable}

\newsymbolglos{count-mc}{n_{\mathrm{MC}}}{Number of Monte Carlo runs}
\newsymbolglos{count-theta}{n_{\theta}}{Number of parameters}
\newsymbolglos{count-window}{\nWind}{Bandwidth of local window}
\newsymbolglos{count-bw}{n_{\mathrm{BW}}}{Number of excited lines in the $3\unit{dB}$ bandwidth of a resonance}
\newsymbolglos{count-order}{\order{A},\order{B},\order{C}}{Order of local models $\LocalModel{A}, \LocalModel{B},\LocalModel{C}$}

@@

\newacronym{EIV}{EIV}{errors-in-variables}
\newacronym{ERNSI}{ERNSI}{European Research Network on System Identification}
\newacronym{ETFE}{ETFE}{empirical transfer function estimate}
